Fail Faster

### 2025-03-19 paper/paper.tex
scala barchart

These results show that staging and \rand choice serve distinct but complementary roles in generator performance. Sampling-heavy generators see greater improvements from faster randomness libraries, while combinator-heavy ones benefit more from staging. Since both factors influence performance in different ways, generator libraries should use both to be performant over diverse workloads.
\subsubsection{Scala} \tr{TODO!!!!}
\begin{figure}[h]
  % \includegraphics[scale=0.33]{allocs.png}
  \includegraphics[scale=0.45]{time.png}
  \caption{Time to generate a value using each strategy, averaged over 10,000 trials. BQ is \bq; Staged SR is \name using \bq's \texttt{splittable\_random} RNG; Staged CSR is \name using \secref{subsection:faster-rng}'s optimized SplitMix. Performance benefits scale with generation size; lower is better.}
  \label{fig:time}
\end{figure}

\begin{figure}[h]
  \centering
  \vspace{0.5em} % Kept the vertical spacing
  \begin{tabular}{lrl}
  \toprule
  Benchmark & \multicolumn{2}{c}{avg. time (ns/op) $\pm$ std. dev.} \\
  \midrule
  BST & $3,560.813$ & $\pm\>1,511.201$ \\
  BST, Staged & $751.287$ & $\pm\>39.890$ \\
  BST (Type-Derived) & $44,619.863$ & $\pm\>5,761.568$ \\
  BST (Type-Derived), Staged & $6,863.209$ & $\pm\>416.188$ \\
  RBT (Type-Derived) & $60,559.846$ & $\pm\>4,901.980$ \\
  RBT (Type-Derived), Staged & $8,511.528$ & $\pm\>324.615$ \\
  STLC & $44,497.623$ & $\pm\>1,970.484$ \\
  STLC, Staged & $7,151.259$ & $\pm\>355.639$ \\
  \bottomrule
  \end{tabular}
 \caption{Benchmarks reimplemented in ScalaCheck show similar performance improvements to \name.}
 \label{fig:scala}
 \end{figure}

>>>>

These results show that staging and \rand choice serve distinct but complementary roles in generator performance. Sampling-heavy generators see greater improvements from faster randomness libraries, while combinator-heavy ones benefit more from staging. Since both factors influence performance in different ways, generator libraries should use both to be performant over diverse workloads.
\subsubsection{Scala}
To confirm that the benefit of staging extends to other strict functional languages, we apply it to ScalaCheck, a widely-used PBT library written in Scala. We implement a subset of our strategies from \secref{subsection:benchmarking} in both staged and unstaged ScalaCheck.

\begin{figure}[h]
  % \includegraphics[scale=0.33]{allocs.png}
  \includegraphics[scale=0.45]{time.png}
  \caption{Time to generate a value using each strategy, averaged over 10,000 trials. BQ is \bq; Staged SR is \name using \bq's \texttt{splittable\_random} RNG; Staged CSR is \name using \secref{subsection:faster-rng}'s optimized SplitMix. Performance benefits scale with generation size; lower is better.}
  \label{fig:time}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.32]{scala.png}
 \caption{Staging in ScalaCheck shows greater speedups than in \bq. SC is ScalaCheck; BQ is \bq.}
 \label{fig:scala}
 \end{figure}
